Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-09-30 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:09:30	21:24:30	(Wed30.Sep)

Throughout this article we have used LCA with variations of the Kaiser window function defined in (\ref{kaiser_window_element}), steered as described in (\ref{steered_window}). However, as shown in \Fig{image_quality} we can obtain similar performance using the trigonometric window function defined in (\ref{trig_window_function} ):
%
\begin{align}
f_m(\alpha) = e^{j\theta_\phi[m]}\Big[\alpha - (1-\alpha)\cos\Big(\frac{2\pi m}{M-1}\Big)\Big]\label{trig_window_function}
\end{align}
%
where $\theta_\phi[m] = \frac{2\pi dm}{\lambda}\sin\phi$ is the $m$th component of the steering vector defined in (\ref{steering_vector}). For this window there is an analytic solution for the optimal value of $\alpha$ in the minimum variance sense. To derive it, we start by inserting the window function (\ref{trig_window_function}) into the beamformer equation (\ref{z}):
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} e^{j\theta_\phi[m]}\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha \a\T\x - (1-\alpha)\b\T\x
\end{align}
%
where $\a$ is the steering vector defined in (\ref{steering_vector}), and 
%
\begin{align}
\b = \text{diag}(\a)\cdot\begin{bmatrix}
     1 &
     \cos\left(\frac{2\pi}{M-1}\right) &
     \cos\left(\frac{2\pi 2}{M-1}\right) &
     \hdots &
     1
     \end{bmatrix}\T.
\end{align}
%
To attain unity gain in the look direction the weights must sum to one:
%
\begin{align}
\w\T\1 &= \big(\alpha\a - (1-\alpha)\b\big)\T\1 = 1
\end{align}
%
This is true for any value of $\alpha$ if $\a\T\1 = 1$ and $\b\T\1 = 1$. Hence, we can preserve the unity gain constraint as long as we normalize $\a$ and $\b$.

Now let $a=\a\T x$ and $b=\b\T x$. The beamformer output can then be written as:
%
\begin{align}
|\w\H\x|^2 &= \Big| \alpha a + (1-\alpha)b \Big|^2 \nn
&= \Big[\alpha a + (1-\alpha)b\Big] \Big[\alpha a^* + (1-\alpha)b^*\Big] \nn
&= \alpha^2 aa^* + \alpha a(1-\alpha)b^* + \alpha a^*(1-\alpha)b + (1-\alpha)^2bb^* \nn
&= \alpha^2 aa^* + \alpha(1-\alpha)(ab^* + a^*b) + (1-\alpha)^2bb^* \nn
&= \alpha^2 (aa^* - (ab^* + a^*b) - bb^*) + \alpha(ab^* + a^*b) + bb^* \nn
\end{align}
%
This is a convex function with a single minimum, which we find by differentiating with respect to $\alpha$ and setting equal to 0:
%
\begin{align}
\frac{\partial}{\partial\alpha} |\w\H\x|^2 
&= 2\alpha (aa^* - (ab^* + a^*b) - bb^*) + ab^* + a^*b = 0,
\end{align}
%
which has the solution:
%
\begin{align}
\alpha &= \frac{ab^* + a^*b}{2(aa^* - (ab^* + a^*b) - bb^*)}.
\end{align}
%
In this computation there are only 4 unique additions and multiplications. If we used this to analytically solve for $\alpha$, but perform the search for $\phi$, the computational complexity of LCA would be of O($8MN_\phi)$ instead of O($MN_\alpha N_\phi)$. The value for $\alpha$ would also be the optimal in the minimum variance sense.

>>>>

Throughout this article we have used LCA with variations of the Kaiser window function defined in (\ref{kaiser_window_element}), steered as described in (\ref{steered_window}). However, as shown in \Fig{image_quality} we can obtain similar performance using the trigonometric window function defined in (\ref{trig_window_function} ):
%
\begin{align}
f_m(\alpha) = e^{j\theta_\phi[m]}\Big[\alpha - (1-\alpha)\cos\Big(\frac{2\pi m}{M-1}\Big)\Big]\label{trig_window_function}
\end{align}
%
where $\theta_\phi[m] = \frac{2\pi dm}{\lambda}\sin\phi$ is the $m$th component of the steering vector defined in (\ref{steering_vector}). For this window there is an analytic solution for the optimal value of $\alpha$ in the minimum variance sense. To derive it, we start by inserting the window function (\ref{trig_window_function}) into the beamformer equation (\ref{z}):
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} e^{j\theta_\phi[m]}\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha \a_\phi\T\x - (1-\alpha)\b_\phi\T\x
\end{align}
%
where $\a_\phi$ is the steering vector defined in (\ref{steering_vector}), and 
%
\begin{align}
\b_\phi = \text{diag}(\a_\phi)\cdot\begin{bmatrix}
     1 &
     \cos\left(\frac{2\pi}{M-1}\right) &
     \cos\left(\frac{2\pi 2}{M-1}\right) &
     \hdots &
     1
     \end{bmatrix}\T.
\end{align}
%
Unity gain in the look direction is ensured as long as the weights sum to one:
%
\begin{align}
\w\T\1 &= \big(\alpha\a_\phi - (1-\alpha)\b_\phi\big)\T\1 = 1.
\end{align}
%
This is true for any value of $\alpha$ if $\a\T\1 = 1$ and $\b\T\1 = 1$. Hence, we can preserve the unity gain constraint as long as we normalize $\a$ and $\b$.

Now let $a=\a\T\x$ and $b=\b\T\x$. The beamformer output can then be written as:
%
\begin{align}
|\w\H\x|^2 &= \Big| \alpha a - (1-\alpha)b \Big|^2 \nn
&= \alpha^2 (aa^* - (ab^* + a^*b) - bb^*) + \alpha(ab^* + a^*b) + bb^*
\end{align}
%
This is a convex function with a single minimum, which we find by differentiating with respect to $\alpha$ and setting equal to 0:
%
\begin{align}
\frac{\partial}{\partial\alpha} |\w\H\x|^2 
&= 2\alpha (aa^* - (ab^* + a^*b) - bb^*) + ab^* + a^*b = 0,
\end{align}
%
which has the solution:
%
\begin{align}
\alpha &= \frac{ab^* + a^*b}{2(aa^* - (ab^* + a^*b) - bb^*)}.
\end{align}
%
In this computation there are only 4 unique additions and multiplications. If we used this to analytically solve for $\alpha$, but perform the search for $\phi$, the computational complexity of LCA would be of O($8MN_\phi)$ instead of O($MN_\alpha N_\phi)$. The value for $\alpha$ would also be the optimal in the minimum variance sense.
